\documentclass{article}
\usepackage{hyperref}

\begin{document}


\title{Erlang UvA Summer School}
\author{Arjan Scherpenisse}

\maketitle
\vspace*{2mm}

\section{A simple chat system}

Use {\tt spawn/1} and {\tt register/2} to start a named server process
which maintains a list of connected clients in its state.  Every
message that the server received, needs to be sent to every other
client. Using {\tt register/2} you can give the server process a name,
so each client  connect to it. The client prints every received
message.

You should create two modules, one for the client and one for the
server; e.g. {\tt chat\_client} and {\tt chat\_server}.

{\tt chat\_server} should export a function {\tt start/0} which spawns
the server process and registers it.

{\tt chat\_client} should export functions {\tt connect/0} to connect
to the server, and {\tt message/1} to send a message to all connected
clients.

Launch different terminals with different Erlang shells to test your
code: one terminal for the server, and one for each client. Use {\tt
  "erl -sname <nodename>"} to start the erlang shells so the erlang
processes are connected. Use the syntax {\tt \{Processname, NodeName\}
  ! Message} to send a message between nodes. For instance: {\tt
  \{server, server@localhost\} ! "foo"} to send {\tt "foo"} to the
registered process called {\tt server} in the {\tt "erl -sname
  server"} terminal.

You can use {\tt net\_adm:ping(nodename@localhost)} to test if your
terminal is connected to the other nodes.

\section{Possible extensions}

\begin{enumerate}

\item Before sending, echo the message locally using {\tt
  io:format/2}. Then, send the message to every other client {\bf but}
  the sending client, to disable local echo.

\item Require a nickname which clients have to enter when
  registering/connecting. Send this nickname along while broadcasting
  client's messages.

\item In the server, maintain a history of the last $N$ messages (with
  timestamps!) that have been sent, and send these upon establishing
  connection.

\item Implement the concept of {\em rooms}: clients can join specific
  rooms and only receive messages sent to those rooms. Find a nice way
  to maintain the list of rooms and clients on the server.

\item Use the \href{https://github.com/basho/rebar}{rebar} build
  system to build your app instead of calling {\tt c(filename).}, and
  use the OTP guidelines to make a proper {\tt gen\_server} out of the
  chat server.
  
\item For more chat inspiration: look at the
  \href{http://tools.ietf.org/html/rfc1459#section-3}{features of the
    IRC protocol}, or, even better, the
  \href{http://xmpp.org/extensions/xep-0045.html}{XMPP (Jabber) MUC}
  (Multi User Chat) specification to implement more features in your
  chat system.

\item Make the server/client communicate with eachother over TCP/IP
  like a proper internet server; use the
  \href{http://ftp.csd.uu.se/pub/mirror/erlang/doc/man/gen\_tcp.html}{{\tt
      gen\_tcp}} OTP module. This is a major step: you need to
  implement proper protocol parsing, no more passing around Erlang
  terms between client/server...

\item Implement a web interface for your module, implementing realtime
  web chat using websockets or longpolling. Woah! This sounds more
  complicated than it is. Ask Arjan about
  \href{http://zotonic.com/}{Zotonic}, and/or check out the
  \href{http://code.google.com/p/zchat/}{zchat} module from Google
  code.

 \item For extra brownie points, document your code, put it in a DVCS,
   publish it on Github, tweet about it, et cetera :-)

\end{enumerate}

\end{document}
